---
title: "Mixed Module Study"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
params:
  echo: no
  rootdir: ~/founder_diet_study
---

Goal: Study mixed module

Idea: Use shinyContrastModule as template, but with mixed WGCNA instead of contrast WGCNA.
Can do same look at strain contrasts, and also use stats summaries.
Make this a new panel for liver app.
Modules will not be different by sex (M,F,B,C) but can do viz by those;
that is, not 4 sets of WGCNA but 1.
Other difference: mixed WGCNA includes multiple datasets;
want to make sure those datasets are selected.

Done:

- Lipid data has been updated with Holland material.
- traitX.rds files updated on ResearchDrive
- `trait` names modified in MixMod$modules; added `dataset` column
- *** need to add column for missing, and keep names of `trait` and `dataset`
- MixMod added to `traitModule.rds`
- plot of Stats for module working in Stats panel
- contrasts computed for eigens as `MixModCond` below
- traits for module identified below

To Do:

- plot of Contrasts for modules and for traits within modules
- add % dropped as characteristic of module
- other summaries (including min `p.value` for `term` within module)
- redeploy both web apps with updated Lipid data
- fix up app.. and shiny.. to work with data

Other change to app: customSetting for default stats (strain:diet, or strain if no diet/condition).

# Setup

```{r}
rawdir <- file.path(params$rootdir, "RawData")
(harmonizeddir <- file.path(params$rootdir, "HarmonizedData"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = harmonizeddir)
```

```{r}
library(foundr)
library(tidyverse)
library(modulr)
```

## Annotation

The experiment annotation file relates the `number` (identifier for `animal`) to the `sex` and `diet`. Needed for `LivRna`.

```{r}
annotfile <- file.path(params$rootdir, "RawData", "Annotation",
                       "mouse annotations for founder diet study.xlsx")
annot <- readxl::read_excel(annotfile) %>%
  mutate(diet = ifelse(as.character(diet_no) == "200339", "HC_LF", "HF_LC"))
```

# Mark's WGCNA Mixed Module

```{r}
mod <- "MixMod"
modparams <- list(
        power = 12, 
        signType = "unsigned",
        minSize = 4)
paramcode <- modparams
if(is.null(paramcode$signType)) {
  paramcode$signType <- "U"
} else {
  paramcode$signType <- ifelse(paramcode$signType == "unsigned", "U", "S")
}
paramcode <- paste(paramcode, collapse = "")
filename <- file.path(mod, paste0(mod, "Module_", paramcode, ".rds"))
```

The object `mixbod` is returned from the call to `wgcna_harmonize()`.
The side effect is to create directory `r mod` in the `harmonizeddir` with `RDS` file
`r filename`.

```{r}
if(file.exists(filename)) {
  mixmod <- readRDS(filename)
} else {
  # This creates the object and saves as RDS.
  moddir <- file.path(params$rootdir, "Primary data from Mark")
  modRdata <- "WGCNA_objects_ms10_mixedmodules_isoforms_lipids_metabolites_founderliver.Rdata"
  mixmod <- modulr::wgcna_harmonize(
    mod,
    moddir,
    modRdata,
    modparams,
    annot,
    harmonizeddir)
  saveRDS(mixmod, filename)
}
```

MixMod trait names, .

```{r}
MixModmodules <- mixmod$value$modules
mixmodtraits <- 
  # Remove ending trait ID.
  str_remove(MixModmodules$trait, "_[0-9]+$")
mixmodLivIso <- str_detect(mixmodtraits, "^ENSMUST0")
mixmodtraits <- mixmodtraits |>
  # Rename isoforms from `ENSMUST0+nnn.aaa` to `aaa_nnn`.
  str_replace("^ENSMUST0+([0-9]+)+\\.(.*)$", "\\2_\\1") |>
  # If name starts with `_` then isoform name is unknown (`NA`).
  str_replace("^_", "NA_")
```

```{r}
MixModmodules$renameTrait <- mixmodtraits
MixModmodules$dataset <- "Unknown"
```

## Examine Modules

```{r}
summary(mixmod)
```

Plot uses `modulr::ggplot_listof_wgcnaModules` through `autoplot` method.

```{r}
ggplot2::autoplot(mixmod)
```

## Traits Matchup

Goal in this section is to match up trait names used by Mark and trait names from harmonized data used in apps. Some notes:

- some traits removed due to
  + too few (<6) strains with data
  + any missing combination of strain*diet*sex
  + possible other reasons not tracked down?
- need for name wrangling due to special characters and Ensemble naming

Traits ingested into Shiny app:

```{r}
if(file.exists(filename <- "liverTraits.rds")) {
  shinytraits <- readRDS(filename)
} else {
  shinytraits <- distinct(readRDS("liverStats.rds"), dataset, trait)
  saveRDS(shinytraits, filename)  
}
```

```{r}
select(
  pivot_wider(
    count(shinytraits, dataset),
    names_from = "dataset", values_from = "n"),
  -MixMod, -LivRna)
```

### Lipids

```{r}
lipid <- filter(shinytraits, dataset == "Lipid")$trait
```

Matches and no-matches.

```{r}
mlipid <- match(make.names(lipid, allow_ = FALSE), make.names(mixmodtraits), nomatch = 0)
c(match = sum(mlipid > 0), nomatch = sum(mlipid == 0))
```

```{r}
MixModmodules$renameTrait[mlipid] <- lipid
MixModmodules$dataset[mlipid] <- "Lipid"
```

## Liver Metabolites

```{r}
livmet <- filter(shinytraits, dataset == "LivMet")$trait
```

```{r}
mlivmet <- match(make.names(livmet, allow_ = FALSE), make.names(mixmodtraits), nomatch = 0)
c(match = sum(mlivmet > 0), nomatch = sum(mlivmet == 0))
```

```{r}
MixModmodules$renameTrait[mlivmet] <- livmet
MixModmodules$dataset[mlivmet] <- "LivMet"
```

## Isoforms

```{r}
liviso <- filter(shinytraits, dataset == "LivIso")$trait
mliviso <- match(make.names(liviso), make.names(mixmodtraits), nomatch = 0)
nliviso <- match(make.names(mixmodtraits), make.names(liviso), nomatch = 0)
(ct <- c(LivIso = sum(nliviso > 0),
  LivIso2 = sum(mliviso > 0),
  NA_iso = sum(nliviso[mixmodLivIso] == 0),
  NA_iso2 = sum(mliviso == 0)))
```

The `r ct["NA_iso"]` `NA_iso` are in the MixMod module but not in the `shinytraits`.
These did not make the cut.

```{r}
MixModmodules$renameTrait[mliviso] <- liviso
MixModmodules$dataset[mliviso] <- "LivIso"
MixModmodules$dataset[mixmodLivIso][nliviso[mixmodLivIso] == 0] <- "LivIsoDrop"
```

Apparently `PI.total`, the remaining "Unknown" is a Lipid trait that was derived after coming from Holland.
It is not in the Holland dataset.

```{r}
MixModmodules <- mutate(MixModmodules,
                        dataset = ifelse(dataset == "Unknown",
                                         "LipidDrop", dataset))
```

```{r}
table(MixModmodules$dataset)
```

# Aligning Mixed Module and Traits

Plan is to modify names in the Mixed Module `mixmod$value$modules` to agree with trait data. Note that with trait as "dataset: trait", some modification of routines
will be necessary.

```{r}
mixmod$value$modules <-
  MixModmodules |>
  mutate(
    trait = renameTrait,
    dropped = (dataset %in% c("LipidDrop","LivIsoDrop")),
    dataset = ifelse(dropped,
                   str_remove(dataset, "Drop"), dataset)) |>
  select(dataset, trait, module, kME, dropped)
```

```{r}
filename <- "traitModule.rds"
traitModule <- readRDS(filename)
```

```{r eval=FALSE}
traitModule$MixMod <- mixmod
saveRDS(traitModule, "traitModule.rds")
```

```{r}
traitModule$MixMod$value$modules |> filter(dropped) |> arrange(dataset)
```

Redo `MixMod` files on ResearchDrive.

```{r eval=FALSE}
links <- read.csv(file.path("..", "RawData", "source.csv"), fill = TRUE) %>%
  mutate(address = file.path(rawdir, address))
harmonize("MixMod", links, foundr::moduleHarmony,
          traitModule["MixMod"],
          condition_name = "diet", normalize = FALSE)
```

See [WGCNA.Rmd](https://github.com/byandell/FounderDietStudy/blob/main/WGCNA.Rmd)
for more detail on module objects used in
[foundr](https://github.com/byandell/foundr) apps.

# Module Eigenvalues

```{r}
MixModSignal <- readRDS(file.path("MixMod", "MixModSignal.rds")) |>
  mutate(dataset = "MixMod")
MixModStats <- readRDS(file.path("MixMod", "MixModStats.rds")) |>
  mutate(dataset = "MixMod")
```

```{r}
MixModCond <- conditionContrasts(MixModSignal, MixModStats, 
        termname = "strain:diet", rawStats = MixModStats)
```

It seems that `eigen_df` is the killer here, depending on
a particular form for rownames that is not adhered to.
Will take some rethinking.

Right now, `eigen_df` is set up for `c("dataset", "strain", "sex")`,
which is fine for contrasts. It needs to be set up for
`c("strain","sex","condition","animal")`
with special handling of `_` in condition.
This is done in another routine.

Also, do not want to paste `sexes` and `module` to expand `module`.

Goal is to have object from eigen_contrast with columns
"dataset" "strain"  "sex"     "module"  "value"   "kME"     "p.value" "size"    "trait"
Only have a subset of these to begin; others come from `module_info().
Look at what is done for contrast.
may need a whole new routine for imported modules.
Rather than having modules constructed by sex, they are constructed across datasets.
So need to capture dataset and not repeat on sex.

Maybe this is not what we want. We have the `eigens` in MixModCond
across the contrasts. What we want to do is match up these with
MixModSignal and get stats from MixModStats for the eigens.
For the traits, we want to use `modules` to match up `dataset` and `trait`
from traitCond with the module.

```{r}
MixModStats |> count(trait) |> count(n, name = "samples")
```

```{r}
MixModSignal |> count(trait) |> count(n, name = "samples")
```

```{r}
MixModCond |> count(trait) |> count(n, name = "samples")
```

```{r}
names(MixModStats)
names(MixModSignal)
names(MixModCond)
```

```{r}
MixModCond
```

The `eigen_contrast` left-joins the 

```{r}
MixModInfo <-
    dplyr::ungroup(
      dplyr::summarize(
        dplyr::group_by(
          mixmod$value$modules,
          .data$module),
        kME = signif(max(abs(kME), na.rm = TRUE), 4),
#        p.value = signif(min(p.value, na.rm = TRUE), 4),
        size = dplyr::n(),
        .groups = "drop"))
MixModInfo
```

```{r}
(object <- 
    dplyr::mutate(
      dplyr::left_join(
        dplyr::rename(MixModCond, module = "trait"),
        # Get information by module.
        MixModInfo,
        by = "module"),
      trait = factor(.data$module, unique(.data$module)),
      module = match(.data$trait, levels(.data$trait))))
```


```{r}
MixModEigen <- foundr:::eigen_contrast_dataset(
  traitModule["MixMod"], MixModCond)
```

# Traits in Modules

```{r}
MixDataTraits <- unite(mixmod$value$modules |> filter(!dropped), 
                       datatrait, dataset, trait, sep = ": ")
```

```{r}
traitStats <- readRDS("liverStats.rds") |>
  unite(datatrait, dataset, trait, sep = ": ", remove = FALSE) |>
  filter(datatrait %in% MixDataTraits$datatrait) |>
  select(-datatrait)
```

```{r}
table(distinct(traitStats, dataset, trait)$dataset)
```

Pick out traits for module `thistle4`

```{r}
readRDS("liverStats.rds") |>
  unite(datatrait, dataset, trait, sep = ": ", remove = FALSE) |>
  filter(datatrait %in%
           filter(MixDataTraits, module == "thistle4")$datatrait) |>
  select(-datatrait) %>%
  filter(term == "strain:diet")
```

# Modules as Traits

```{r}
PhysioSignal <- readRDS(file.path("Physio", "PhysioSignal.rds")) |>
  mutate(dataset = "Physio")
PhysioStats <- readRDS(file.path("Physio", "PhysioStats.rds")) |>
  mutate(dataset = "Physio")
```

```{r}
traitModule$Physio$Female$modules
```

```{r}
PhysioCond <- conditionContrasts(PhysioSignal, PhysioStats, 
        termname = "strain:diet", rawStats = PhysioStats)
PhysioEigen <- foundr:::eigen_contrast_dataset(
  traitModule["Physio"], PhysioCond)
```

```{r}
PhysioCond
```

```{r}
PhysioEigen
```

```{r}
knitr::knit_exit()
```

## Further checking of Liver Isoforms

This is not pursued further, but provided justification of the dropped isoforms.
Basically, these did not have sufficient data to estimate stats. For some,
there were only female entries. Some of this code may not work at this point.

Further checking of mismatch. For the `r sum(mliviso == 0)` isoforms that did not match exactly,
none of their gene names or isoform IDs match. These will be excluded.

```{r}
c(gene = sum(is.na(match(str_remove(liviso[mliviso == 0], "_[0-9]+$"), 
                         mixmodtraits[mixmodLivIso]))),
  ID = sum(is.na(match(str_remove(liviso[mliviso == 0], "^.*_"), 
                       mixmodtraits[mixmodLivIso]))))
```

For the `r sum(nliviso[mixmodLivIso] == 0)` that are not included in the
shiny app, most of them do not have sufficient data for the following reasons:

- too few (<6) strains with data
- any missing combination of strain*diet*sex
- possible other reasons not tracked down?

Below we characterize these dropped isoforms.

```{r}
missIso <- mixmodtraits[mixmodLivIso][nliviso[mixmodLivIso] == 0]
# LivIso statistics for all traits
LivIsoStats <- readRDS("~/Documents/Research/attie_alan/FounderDietStudy/data/LivIsoStats.rds")
# Drop trait if missing p-value for any combination of `strain*diet*sex`
dropTraits <- unique(
  dplyr::filter(
    # Ignore "noise" term as it has no p.value.
    dplyr::filter(
      LivIsoStats,
      !(.data$term %in% c("noise", "rawSD"))),
    is.na(.data$p.value))$trait)
md <- match(missIso, dropTraits)
md_na <- sum(is.na(md))
c(dropped = length(dropTraits),
  notmixmod = length(dropTraits) - length(md),
  mixmoddrop = length(md) - md_na,
  nostats = md_na)
head(missIso[is.na(md)])
```

There were `r md_na` traits that were dropped before they got to the stats calculations.
Let's look at those.

```{r}
LivIsoData <- readRDS("~/Documents/Research/attie_alan/FounderDietStudy/data/LivIsoData.rds")
```

```{r}
LivIsoTraits <- LivIsoData |>
  filter(trait %in% missIso[is.na(md)]) |>
  distinct(trait, strain) |>
  count(trait)
LivIsoTraits |>
  count(n, name = "nstrain") |>
  pivot_wider(names_from = "n", values_from = "nstrain")
```

The `r sum(LivIsoTraits$n < 6)` that have fewer than 6 strains were dropped.
That leaves the following `r md_na - sum(LivIsoTraits$n < 6)` traits:

```{r}
LivIsoTraits |>
  filter(n >= 6) |>
  arrange(n, trait)
```

These traits have 24 or more measurements and 6 or more strains.

```{r}
LivIsoData <- filter(LivIsoData, trait %in% filter(LivIsoTraits, n >= 6)$trait)
stem((LivIsoData |>
  count(trait))$n)
```

All of these traits only have females. They fail the model fit of `lm` due
to failure in `model.matrix`. That is, we need 2 levels over every factor
and factor combination to successfully run analysis. Here are the the
counts by factors and factor combinations. The contrasts for `model.matrix`
used by `lm` are even more stringent.

```{r}
ctfn <- function(x, factors) count(distinct(select(x, any_of(factors))))$n
purfn <- function(x, factors) unlist(purrr::map(split(x, x$trait), ctfn, factors))
data.frame(
  # Single factors
  strain = purfn(LivIsoData, c("strain")),
  sex = purfn(LivIsoData, c("sex")),
  diet = purfn(LivIsoData, c("condition")),
  # Pairs of factors
  strain_sex = purfn(LivIsoData, c("strain", "sex")),
  strain_diet = purfn(LivIsoData, c("strain", "condition")),
  sex_diet = purfn(LivIsoData, c("diet", "sex")),
  strain_sex_diet = purfn(LivIsoData, c("strain", "sex", "condition")))
```

```{r}
liv
```

```{r}
MixModEigen <- foundr::eigen_contrast(mixmod$value, )
```

```{r}
summary(LivMetEigen)
```

