---
title: "Module Visualization"
author: "Brian Yandell"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
runtime: shiny
params:
  echo: no
  rootdir: ~/founder_diet_study
---

```{r}
rawdir <- file.path(params$rootdir, "RawData")
harmonizeddir <- file.path(params$rootdir, "HarmonizedData")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = params$echo, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = harmonizeddir)
```

```{r}
getwd()
```

```{r}
datadir <- "."
datadir <- file.path(datadir)
```

```{r}
devtools::install_cran("plotly") #  not yet on UW dataviz
devtools::install_cran("markdown") #  not yet on UW dataviz
devtools::install_cran("cowplot") #  not yet on UW dataviz
devtools::install_cran("patchwork") #  not yet on UW dataviz
devtools::install_cran("ggdendro") # not yet in UW dataviz
devtools::install_github("byandell/foundr")
```

This document selects a `dataset` and examines WGCNA modules. Selecting a `Sex Type` enables visualization of stats for the modules as well as the WGCNA dendogram and color panel. 

```{r}
#datasets <- c("LivMet","PlaMet0","PlaMet120","Metab")
datasets <- "LivMet"
sexes <- c("F","M","F+M","F-M")
names(sexes) <- c("Female","Male","BothSexes","SexContrast")
responses <- as.vector(sexes)
#responses <- c("value","cellmean","signal","rest","noise")
```

```{r}
traitModule <- readRDS(file.path(datadir, "traitModule.rds"))
```

```{r}
shiny::fluidRow(
  shiny::column(
    6,
    shiny::selectInput(
      "dataset", "Dataset:",
      datasets)),
  shiny::column(
    6,
    shiny::selectInput(
      "response", "Sex Type:",
      as.vector(sexes)))
)
```

# Module Eigentrait Contrasts across Strain and Sex

```{r}
traitSignal <-
    dplyr::filter(
      readRDS(file.path(datadir, "traitSignal.rds")),
      dataset %in% datasets)
traitStats <-
    dplyr::filter(
      readRDS(file.path(datadir, "traitStats.rds")),
      dataset %in% datasets)
```

```{r}
traitContr <- shiny::reactive({
  dplyr::filter(
    foundr::conditionContrasts(
      dplyr::filter(traitSignal,
                    .data$dataset %in% shiny::req(input$dataset)),
      dplyr::filter(traitStats,
                    .data$dataset %in% shiny::req(input$dataset)),
      termname = "signal"),
    .data$p.value <= 0.05)
})
```

```{r}
traitEigen <- shiny::reactive({
  shiny::req(input$dataset)
  
  modulr::eigen_contrast(traitModule[[input$dataset]],
                              traitContr())
})
```
  
```{r}
shiny::renderPlot({
  plot(traitEigen(), bysex = shiny::req(input$response))
})
```

```{r}
shiny::renderPlot({
  plot(traitEigen(), bysex = shiny::req(input$response),
       volcano = TRUE,
       threshold = c(SD = 0.5, p = 0.1))
})
```

```{r}
DT::renderDataTable({
  shiny::req(traitEigen())
  
  summary(traitEigen())
})
```

# Relating Traits to Eigentraits in Modules

```{r}
module_choices <- shiny::reactive({
  unique(traitModule[[shiny::req(input$dataset)]][[shiny::req(input$response)]]$modules$module)
})
```

```{r}
shiny::renderText(module_choices())
```

```{r}
shiny::renderUI({
  shiny::selectInput("module", "Module:", shiny::req(module_choices()))
})
```

```{r}
eigenComp <- shiny::reactive({
  modulr::eigen_traits(traitModule[[shiny::req(input$dataset)]],
                          sexname = shiny::req(input$response),
                          modulename = shiny::req(input$module),
                          traitContr())
})
```

```{r}
shiny::renderPlot({
  plot(eigenComp(), bysex = shiny::req(input$response))
})
```

```{r}
shiny::renderPlot({
  plot(eigenComp(), bysex = shiny::req(input$response), volcano = TRUE,
       threshold = c(SD = 1, p = 0.2),
       ylab("eigen -log10(1-|kME|)"))
})
```

```{r}
DT::renderDataTable({
  shiny::req(eigenComp())
  
  summary(eigenComp())
})
```


# WGCNA Dendrogram

```{r}
shiny::renderPlot({
  shiny::req(input$dataset, input$response)
  foundr::ggplot_listof_wgcnaModules(traitModule[[input$dataset]],
                                     input$response)
})
```

```{r}
DT::renderDataTable({
  shiny::req(input$dataset, input$response)
  DT::datatable(
    dplyr::filter(
      summary(traitModule[[input$dataset]]),
      response == input$response),
    options = list(paging =TRUE, pageLength = 5))
})
```

# Planned Updates

- order eigen_contrast sex for summary
- set alpha close to one for conditionContrasts
- separate these two parts by buttons
- check on division by rawSD
- titles and explanations
- default choices for second part
- separate method? show kME values, less offset for module, etc.
